Parsing Postfix log files

### 2009-03-09 doc/logparser-architecture.tex
Spell check.

    \item [Classify the input]  The pseudo-code above shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or prioritisations
        (\sectionref{rules in architecture}) that complicate the process.

@@

Once conditions pass a certain level of complexity determining whether two
conditions overlap becomes intractable, because it requires so much
knowledge of other rules and possibly even actions.  E.g.\ given two rules
with conditions \verb!verbose == true! and \verb!silent == true!, logically
these rules should not overlap, yet there is nothing to stop both variables
being set to true by one or more rules or actions.  If variables used in
conditions can be set by actions, determining whether or not two conditions
overlap is impossible: the Halting Problem show that it is impossible for
one program to answer questions about another program's behaviour if the
inquiring program is implemented on a compuational machine of the equal or
lesser power to the program being inquired into.

>>>>

    \item [Classify the input]  The pseudo-code above shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or rule priorities
        (\sectionref{rules in architecture}) that complicate the process.

@@

Once conditions pass a certain level of complexity determining whether two
conditions overlap becomes intractable, because it requires so much
knowledge of other rules and possibly even actions.  E.g.\ given two rules
with conditions \verb!verbose == true! and \verb!silent == true!, logically
these rules should not overlap, yet there is nothing to stop both variables
being set to true by one or more rules or actions.  If variables used in
conditions can be set by actions, determining whether or not two conditions
overlap is impossible: the Halting Problem show that it is impossible for
one program to answer questions about another program's behaviour if the
inquiring program is implemented on a computational machine of the equal or
lesser power to the program being inquired into.

### 2009-03-09 doc/logparser-conclusion.tex
Divided rules section in architecture into subsections.
Spell check.

@@

The architecture not only substantially eases the parsing of new log lines,
it makes adding new actions (\sectionref{adding new actions in
implementation}) a relatively simple task.  The simplicity of adding a new
action frees the implementor from worrying about how their action might
interfere with existing parsing, so they can focus on correctly
implementing their new action.

>>>>

XXX PULL CONTENT FROM THE PAPER\@.

@@

The architecture not only substantially eases the parsing of new log lines,
it makes adding new actions (\sectionref{adding new actions in
implementation}) a relatively simple task.  The simplicity of adding a new
action frees the implementer from worrying about how their action might
interfere with existing parsing, so they can focus on correctly
implementing their new action.

### 2009-03-09 doc/logparser-implementation.tex
Spell check.

        In the original algorithm the purpose of the preprocessing stage
        was to reduce the memory consumption of the program.  In the first
        pass it generated a hash~\cite{hash-functions}\glsadd{hash} from a
        small range of values for each word of each log line, incrementing
        a counter for each hash.  The counters will be used in the next
        pass to filter out words: if the word's hash does not have a high
        frequency, the word itself cannot have a high frequency, so there
        is no need to maintain a counter for it, reducing the number of
        counters required and thus the program's memory consumption.

@@

Adding new actions is not as easy as adding new rules, though care has been
taken in the parser's design and implementation to make adding new actions
as painless as possible.  The implementor writes a subroutine that accepts
the standard arguments given to actions, and registers it as an action by
calling the framework subroutine add\_actions() with the name of the new
action as a parameter.  The new action must be registered before the rules
are loaded, because it is an error for a rule to specify an unregistered
action; this helps catch mistakes made when adding new rules.  No other
work is required from the implementor to integrate the action into the
parser; all of their attention and effort can be focused on correctly
implementing their action.  The only negative aspect is that the process
involves editing the parser source code, which makes upgrading to a later
version of the parser more difficult, though by no means impossible.  If
the author of the new action wishes, they can take advantage of the
parser's object oriented implementation by subclassing it and implementing
their changes in the derived class, allowing future upgrades of the parser
with greatly reduced chance of conflicts.\footnote{The real difference
between the two approaches is where the new code is placed.  The simpler
option is to change the parser code directly, but those changes will then
have to be made to subsequent versions of the parser, and as the scope of
the changes increases so does the chance of conflict, or mistakes when
copying the action.  The more time consuming option is to write a subclass
containing the new actions and change the program that invokes the parser
so that it uses the subclass rather than the parser; the changes required
to the program invoking the parser are minor and much less likely to lead
to conflicts when upgrading to future versions of the parser.  An
alternative is to submit new actions to the author of the parser for
inclusion in future versions, resulting in two benefits: the new actions do
not need to be maintained separately, and other users of the parser can
avail of the new functionality.} The action may need to extend the list of
valid combinations described in \sectionref{out of order log lines} if the
new action creates a different set of acceptable programs, but this is
unlikely to occur, as it would require parsing log lines from Postfix
components the parser currently ignores.\footnote{The mail server used for
development does not utilise either the \daemon{lmtp} or \daemon{virtual}
delivery agents, so this parser does not have rules to handle log lines
from those components.  Adding new rules to parse those component's log
lines is a simple process, though if their behaviour differs significantly
from the \daemon{smtp} or \daemon{local} delivery agents new actions may be
required.  The mail server in question is a production mail server handling
mail for a university department; the benefit of using this server is that
the log files used exhibit the idiosyncrasies and peculiarities a mail
server in the wild must deal with, but the downside is that significantly
altering the configuration just to log messages from a different Postfix
component is not an option.}

@@

If a mail is mis-classified as a bounce message it will almost certainly
have been generated internally by Postfix; arguably mis-classification in
this case is a benefit rather than a drawback, as other mails generated
internally by Postfix will be handled correctly.  Postfix 2.3 (and
hopefully subsequent versions) log the creation of a bounce message.

>>>>

        In the original algorithm the purpose of the pre-processing stage
        was to reduce the memory consumption of the program.  In the first
        pass it generated a hash~\cite{hash-functions}\glsadd{hash} from a
        small range of values for each word of each log line, incrementing
        a counter for each hash.  The counters will be used in the next
        pass to filter out words: if the word's hash does not have a high
        frequency, the word itself cannot have a high frequency, so there
        is no need to maintain a counter for it, reducing the number of
        counters required and thus the program's memory consumption.

@@

Adding new actions is not as easy as adding new rules, though care has been
taken in the parser's design and implementation to make adding new actions
as painless as possible.  The implementer writes a subroutine that accepts
the standard arguments given to actions, and registers it as an action by
calling the framework subroutine add\_actions() with the name of the new
action as a parameter.  The new action must be registered before the rules
are loaded, because it is an error for a rule to specify an unregistered
action; this helps catch mistakes made when adding new rules.  No other
work is required from the implementer to integrate the action into the
parser; all of their attention and effort can be focused on correctly
implementing their action.  The only negative aspect is that the process
involves editing the parser source code, which makes upgrading to a later
version of the parser more difficult, though by no means impossible.  If
the author of the new action wishes, they can take advantage of the
parser's object oriented implementation by sub-classing it and implementing
their changes in the derived class, allowing future upgrades of the parser
with greatly reduced chance of conflicts.\footnote{The real difference
between the two approaches is where the new code is placed.  The simpler
option is to change the parser code directly, but those changes will then
have to be made to subsequent versions of the parser, and as the scope of
the changes increases so does the chance of conflict, or mistakes when
copying the action.  The more time consuming option is to write a subclass
containing the new actions and change the program that invokes the parser
so that it uses the subclass rather than the parser; the changes required
to the program invoking the parser are minor and much less likely to lead
to conflicts when upgrading to future versions of the parser.  An
alternative is to submit new actions to the author of the parser for
inclusion in future versions, resulting in two benefits: the new actions do
not need to be maintained separately, and other users of the parser can
avail of the new functionality.} The action may need to extend the list of
valid combinations described in \sectionref{out of order log lines} if the
new action creates a different set of acceptable programs, but this is
unlikely to occur, as it would require parsing log lines from Postfix
components the parser currently ignores.\footnote{The mail server used for
development does not utilise either the \daemon{lmtp} or \daemon{virtual}
delivery agents, so this parser does not have rules to handle log lines
from those components.  Adding new rules to parse those component's log
lines is a simple process, though if their behaviour differs significantly
from the \daemon{smtp} or \daemon{local} delivery agents new actions may be
required.  The mail server in question is a production mail server handling
mail for a university department; the benefit of using this server is that
the log files used exhibit the idiosyncrasies and peculiarities a mail
server in the wild must deal with, but the downside is that significantly
altering the configuration just to log messages from a different Postfix
component is not an option.}

@@

If a mail is misclassified as a bounce message it will almost certainly
have been generated internally by Postfix; arguably misclassification in
this case is a benefit rather than a drawback, as other mails generated
internally by Postfix will be handled correctly.  Postfix 2.3 (and
hopefully subsequent versions) log the creation of a bounce message.

### 2009-03-09 doc/logparser-other-parsers-review.tex
Spell check.

\acronym{LMA} does not deal with any of the complications discussed in
\sectionref{complications}, except for correlating log lines by queueid;
not correlating log lines by pid means it cannot correlate most rejections.
It does not differentiate between different types of rejections, so it is
not suitable for the purposes of this project; the data about which
restriction caused the rejection is discarded, whereas the main goal of
this project is to retain that data to aid optimisation and evaluation of
anti-spam techniques.  \acronym{LMA} fails to parse Postfix log files
generated on Solaris hosts because the fields automatically prepended to
each log line differ from those added on Linux hosts; log files from
Solaris hosts (and possibly other operating systems) thus require
preprocessing before parsing by \acronym{LMA}.  The \numberOFlogFILES{}
preprocessed test log files were parsed without complaint by \acronym{LMA},
although it produced 32 entries in the output \acronym{CSV} file for every
rejection in the input log file; it also missed some 40\% of delivered
mail.  Once these deficiencies were discovered the author did not spend any
more time checking the results.

@@

In summary, Sawmill suffers from supporting so many types of log files: it
parses many types of log files, but none of them very well; it is probably
much more useful when parsing log files where each log line is
self-contained (e.g.\ web server log files), rather than log files
containing interlinked log lines.  It is not suitable as a base for this
parser, as the source code made available is obfuscated and not intended
for modification; in addition the architecture would probably need to be
overhauled or replaced to deal with correlating log lines.

@@

        \item [\nialpha{}] Sawmill's source code is available in an
            obfuscated form, so that customers can compile it on platforms
            that pre-compiled binaries are not available for.

>>>>

\acronym{LMA} does not deal with any of the complications discussed in
\sectionref{complications}, except for correlating log lines by queueid;
not correlating log lines by pid means it cannot correlate most rejections.
It does not differentiate between different types of rejections, so it is
not suitable for the purposes of this project; the data about which
restriction caused the rejection is discarded, whereas the main goal of
this project is to retain that data to aid optimisation and evaluation of
anti-spam techniques.  \acronym{LMA} fails to parse Postfix log files
generated on Solaris hosts because the fields automatically prepended to
each log line differ from those added on Linux hosts; log files from
Solaris hosts (and possibly other operating systems) thus require
pre-processing before parsing by \acronym{LMA}.  The \numberOFlogFILES{}
pre-processed test log files were parsed without complaint by
\acronym{LMA}, although it produced 32 entries in the output \acronym{CSV}
file for every rejection in the input log file; it also missed some 40\% of
delivered mail.  Once these deficiencies were discovered the author did not
spend any more time checking the results.

@@

In summary, Sawmill suffers from supporting so many types of log files: it
parses many types of log files, but none of them very well; it is probably
much more useful when parsing log files where each log line is
self-contained (e.g.\ web server log files), rather than log files
containing interlinked log lines.  It is not suitable as a base for this
parser, as the source code made available is encrypted and not intended
for modification; in addition the architecture would probably need to be
overhauled or replaced to deal with correlating log lines.

@@

        \item [\nialpha{}] Sawmill's source code is available in an
            encrypted form, so that customers can compile it on platforms
            that pre-compiled binaries are not available for.

### 2009-03-09 doc/logparser-results.tex
Fixed the problem where \label{} and \ref{} didn't work properly with glossary chapters.
Corrections, improvements, and new content in architecture chapter. Added SGAI and Elsevier entries to bibliography. Some minor cleanup elsewhere.
Some minor formatting fixes. Documented some of 'make check'. Added check for sentences with only one space separating them.
Spell check.

\subsection{Caching Compiled Regexs}

@@

Before the perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and
optimised to improve the speed of matching.  This compilation and
optimisation takes CPU time ---in many cases it takes far more CPU time
than the actual matching.  If the interpreter is certain that a regex
will not change it will automatically cache the compilation results for
later use.   The results of compiling a dynamically generated regex can
be cached and used in preference to the original regex, but it is the
responsibility of the programmer to do this.  The Postfix parser loads the
rules it uses from a database, and thus each rule's regex is compiled
and the result cached to reduce parsing time.

@@

\newpage{} % XXX

@@

XXX REWRITE COVERAGE SUMARY IF NECESSARY\@.

>>>>

\subsection{Caching Compiled Regexes}

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and
optimised to improve the speed of matching.  This compilation and
optimisation takes CPU time ---in many cases it takes far more CPU time
than the actual matching.  If the interpreter is certain that a regex
will not change it will automatically cache the compilation results for
later use.   The results of compiling a dynamically generated regex can
be cached and used in preference to the original regex, but it is the
responsibility of the programmer to do this.  The Postfix parser loads the
rules it uses from a database, and thus each rule's regex is compiled
and the result cached to reduce parsing time.

@@

@@

XXX REWRITE COVERAGE SUMMARY IF NECESSARY\@.
